Calendar-based graphics for visualizing people's daily schedules

### 2017-08-08 main.Rmd
done with construction, some of options

Figure \ref{fig:flinders-2016} shows the line glyphs framed in the monthly calendar over the year 2016. This is achieved by the \code{frame_calendar} function computing the new coordinates according to the input data variables; in turn the rearranged data values are plotted using the \pkg{ggplot2} package [@R-ggplot2], which is an implementation of the grammar of graphics [@wilkinson2006grammar; @wickham2010layered].

@@

In order to make calendar-based graphics more accessible and informative, reference lines dividing each cell and block as well as labels indicating week day and month are also computed before plotting construction.

@@

The month labels located on the top left are obtained through $(\min{(x)}, \max{(y)})$ for every $(m, n)$. The week day texts are uniformly positioned on the bottom of the whole canvas, that is $\min{(y)}$, with the central position of a cell $x / 2$ for each $j$.

@@

There are several options provided for the \code{frame_calendar} function to initialize and adjust the display of a calendar plot:

@@

Assuming that tidy data [@wickham2014tidy] is the underlying data structure, the parameter \code{x} takes a variable that will be mapped to the horizontal axis and the parameter \code{y} mapped to the vertical axis for plot construction. In Figure \ref{fig:flinders-2016}, for example, the \code{x} is the variable specifying the time of the day, and the \code{y} is the variable representing the hourly counts. The \code{date} argument is given by the date variable that determines the correct order of the calendar layout.

The algorithm can be extended to display data from a single month up to a few years. The number of rows and columns in the layout can be specified with the arguments \code{nrow} and \code{ncol}. If the one would like to visualize data spanning over three years, for example, \code{nrow = 3} and \code{ncol = 12} would be an appropriate choice when assessing the differences of a given month across the years.

In Section \ref{sec:algorithm} we only illustrated that grids are laid out horizontally. The vertical direction can be enabled by swapping $i$ and $j$ in equation \eqref{eq:grid}, which occurs when the argument \code{dir} is set to \code{"v"}. This benefits users who are accustomed to calendars that are organized vertically, which is common in some countries. Next we shall describe some of the arguments that allow for alternate displays.

>>>>

Figure \ref{fig:flinders-2016} shows the line glyphs framed in the monthly calendar over the year 2016. This is achieved by the \code{frame_calendar} function, which computes the coordinates on the calendar for the input data variables. These can then be plotted using the usual \pkg{ggplot2} package [@R-ggplot2] functions. All of the grammar of graphics [@wilkinson2006grammar; @wickham2010layered] can be applied. XXX use Hadley's book as the reference rather than the JCGS paper

@@

In order to make calendar-based graphics more accessible and informative, reference lines dividing each cell and block as well as labels indicating week day and month are also computed before plot construction.

@@

The month labels located on the top left using $(\min{(x)}, \max{(y)})$ for every $(m, n)$. The week day texts are uniformly positioned on the bottom of the whole canvas, that is $\min{(y)}$, with the central position of a cell $x / 2$ for each $j$.

@@

The algorithm has several optional parameters that modify the layout, direction of display, scales, white space and switching to polar coordinates. These are accessible to the user by the inputs to the function  \code{frame_calendar}:

@@

It is assumed that the `data` is in tidy format [@wickham2014tidy], and \code{x}, \code{y} are the variables that will be mapped to the horizontal and vertical axes in each cell. For example, the \code{x} is the time of the day, and \code{y} is the count (Figure \ref{fig:flinders-2016}). The \code{date} argument specifies the date variable used to construct the calendar layout.

The algorithm handles displaying a single month or several years. The  arguments \code{nrow} and \code{ncol} specify the layout of multiple months. For some time frames, some arrangements may be more beneficial than others. For example, to display data for three years, setting \code{nrow = 3} and \code{ncol = 12} would show each year on a single row.

By default, grids are laid out horizontally. This can be transposed by setting the \code{dir} parameter to \code{"v"}, in which case $i$ and $j$ are swapped in the equation \eqref{eq:grid}. This can be useful for creating  calendar layouts for countries where vertical layout is the convention.

### 2017-08-08 main.tex
done with construction, some of options

Figure \ref{fig:flinders-2016} shows the line glyphs framed in the
monthly calendar over the year 2016. This is achieved by the
\code{frame_calendar} function computing the new coordinates according
to the input data variables; in turn the rearranged data values are
plotted using the \pkg{ggplot2} package \citep{R-ggplot2}, which is an
implementation of the grammar of graphics
\citep{wilkinson2006grammar, wickham2010layered}.

@@

The month labels located on the top left are obtained through
\((\min{(x)}, \max{(y)})\) for every \((m, n)\). The week day texts are
uniformly positioned on the bottom of the whole canvas, that is
\(\min{(y)}\), with the central position of a cell \(x / 2\) for each
\(j\).

@@

Assuming that tidy data \citep{wickham2014tidy} is the underlying data
structure, the parameter \code{x} takes a variable that will be mapped
to the horizontal axis and the parameter \code{y} mapped to the vertical
axis for plot construction. In Figure \ref{fig:flinders-2016}, for
example, the \code{x} is the variable specifying the time of the day,
and the \code{y} is the variable representing the hourly counts. The
\code{date} argument is given by the date variable that determines the
correct order of the calendar layout.

The algorithm can be extended to display data from a single month up to
a few years. The number of rows and columns in the layout can be
specified with the arguments \code{nrow} and \code{ncol}. If the one
would like to visualize data spanning over three years, for example,
\code{nrow = 3} and \code{ncol = 12} would be an appropriate choice when
assessing the differences of a given month across the years.

In Section \ref{sec:algorithm} we only illustrated that grids are laid
out horizontally. The vertical direction can be enabled by swapping
\(i\) and \(j\) in equation \eqref{eq:grid}, which occurs when the
argument \code{dir} is set to \code{"v"}. This benefits users who are
accustomed to calendars that are organized vertically, which is common
in some countries. Next we shall describe some of the arguments that
allow for alternate displays.

>>>>

Figure \ref{fig:flinders-2016} shows the line glyphs framed in the
monthly calendar over the year 2016. This is achieved by the
\code{frame_calendar} function, which computes the coordinates on the
calendar for the input data variables. These can then be plotted using
the usual \pkg{ggplot2} package \citep{R-ggplot2} functions. All of the
grammar of graphics \citep{wilkinson2006grammar, wickham2010layered} can
be applied. XXX use Hadley's book as the reference rather than the JCGS
paper

@@

The month labels located on the top left using
\((\min{(x)}, \max{(y)})\) for every \((m, n)\). The week day texts are
uniformly positioned on the bottom of the whole canvas, that is
\(\min{(y)}\), with the central position of a cell \(x / 2\) for each
\(j\).

@@

It is assumed that the \texttt{data} is in tidy format
\citep{wickham2014tidy}, and \code{x}, \code{y} are the variables that
will be mapped to the horizontal and vertical axes in each cell. For
example, the \code{x} is the time of the day, and \code{y} is the count
(Figure \ref{fig:flinders-2016}). The \code{date} argument specifies the
date variable used to construct the calendar layout.

The algorithm handles displaying a single month or several years. The
arguments \code{nrow} and \code{ncol} specify the layout of multiple
months. For some time frames, some arrangements may be more beneficial
than others. For example, to display data for three years, setting
\code{nrow = 3} and \code{ncol = 12} would show each year on a single
row.

By default, grids are laid out horizontally. This can be transposed by
setting the \code{dir} parameter to \code{"v"}, in which case \(i\) and
\(j\) are swapped in the equation \eqref{eq:grid}. This can be useful
for creating calendar layouts for countries where vertical layout is the
convention.
